---
title: "Untitled"
output: pdf_document
date: "2023-07-24"
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(fixest)
library(modelsummary)
library(panelsummary)
library(kableExtra)
library(did2s)

rollout_dates <- read_csv("created_data/rollout_dates.csv") %>% 
  mutate(across(c(-1), ~mdy(.)))

if (!exists("dispatch_panel")){
  dispatch_panel <- read_csv(here::here("analysis_data/xxdispatches_clevel.csv"))
  dispatch_panel_p1 <- dispatch_panel %>% 
    filter(priority_code == 1)
}
```

# Notes on ShotSpotter

## ShotSpotter Statistics

```{r}
library(tidyverse)
sst <- read_csv(here::here("created_data/sst_dispatch_cpd.csv"))

```

It appears that 95% of SST dispatches do not result in anything. According to the Office of the Inspector General, "If these activities
produce evidence of a shooting or any other criminal activity, a corresponding criminal incident code will be assigned. If there is no such evidence, then the event will receive a miscellaneous incident disposition code."


```{r}
## fraction of SST alerts that turn up nothing
sst_fraction_found <- sst %>% 
  count(final_disposition_description,rd, sort = T) %>% 
  mutate(total = sum(n)) %>% 
  mutate(n/total) %>%
  head()

sst_fraction_found
```



```{r}
## per day amounts of SST
per_day_sst <- sst %>% 
  mutate(date = as_date(entry_received_date), .before = 1) %>% 
  filter(date >= as_date("2018-05-16")) %>% 
  group_by(date) %>% 
  summarize(sst_per_day = n()) %>% 
  summarize(mean(sst_per_day)) %>% 
  pull() %>% 
  round(2)
```

After the final implementation date of 2018-05-16, there is, on average, `r per_day_sst` SST alerts every day across the city.


```{r}
sst %>% 
  mutate(date = as_date(entry_received_date), .before = 1) %>% 
  mutate(hour = hms::as_hms(entry_received_date), .before = 1) %>% 
  ggplot(aes(hour)) +
  geom_density(fill = "dark red", alpha = 0.5) +
  scale_x_time() +
  theme_minimal()
```

Juneviles account for 11% of victims in our data over 2016-2022. We cannot connect these to 911 calls.


## victimization over time

```{r}


victims_overal <- read_csv(here::here("analysis_data/victims_panel.csv"))

victims_overal %>% 
  mutate(
          year = year(date),
          month = month(date),
          day = day(date),
          year_month = mdy(paste0(month, "-1-", year))) %>% 
  mutate(victim_gun = num_any_victim - num_victim_no_gun) %>% 
  group_by(year_month) %>% 
  summarize(n = sum(victim_gun, na.rm  = T)) %>% 
  ggplot(aes(year_month, n)) +
  geom_line()

```



## Percentage of Priority 1

```{r}

sst <- sst %>%
  mutate(district = parse_number(district)) %>% 
  mutate(date = ymd_hms(entry_received_date) %>% as_date()) %>% 
  left_join(rollout_dates) %>% 
  mutate(treatment = if_else(shotspot_activate <= date, 1, 0), 
         treatment_sdsc = if_else(sdsc <= date, 1, 0 ),
         treatment_official = if_else(shotspot_activate_official <= date, 1, 0),
         treatment_first_shot = if_else(shotspot_activate_first_shot <= date, 1, 0),
         treatment_cpd = if_else(shotspot_activate_cpd <= date, 1, 0),
         .by = district) %>% 
  mutate(never_treated = if_else(is.na(treatment),1, 0), .by = district) %>% 
  mutate(treatment = if_else(is.na(treatment), 0, treatment
  ),
  treatment_sdsc = if_else(is.na(treatment_sdsc), 0 , treatment_sdsc),
  treatment_official = if_else(is.na(treatment_official), 0, treatment_official),
  treatment_first_shot = if_else(is.na(treatment_first_shot), 0, treatment_first_shot),
  .by = district)


sst <- sst %>% 
  mutate(priority_code = parse_number(priority_code))

sst <- sst %>% 
  mutate(sst = 1) 

x <- dispatch_panel_p1 %>%
  bind_rows(sst)

x %>% 
  filter(never_treated == 0) %>% 
  filter(treatment == 1) %>% 
  filter(sst == 1) %>% 
  mutate(sst= replace_na(sst, 0)) %>% 
  count(date) %>% 
  summarize(mean(n))
 

dispatch_panel_p1 %>% 
  filter(date <= as_date("2017-02-01")) %>% 
  filter(final_dispatch_description == "SHOTS FIRED") %>% 
  count(date) %>% 
  summarize(mean(n))
```

